---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gmeans

<!-- badges: start -->
[![R-CMD-check](https://github.com/m-muecke/gmeans/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/m-muecke/gmeans/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/gmeans)](https://CRAN.R-project.org/package=gmeans)
[![R-universe version](https://m-muecke.r-universe.dev/gmeans/badges/version)](https://m-muecke.r-universe.dev/gmeans)
<!-- badges: end -->

The aim of the package is to provide an implementation of the G-means algorithm in R.
The G-means algorithm is a clustering algorithm that extends the k-means algorithm by
automatically determining the number of clusters.
The algorithm was introduced by [Hamerly and Elkan (2003)](https://proceedings.neurips.cc/paper_files/paper/2003/file/234833147b97bb6aed53a8f4f1c7a7d8-Paper.pdf).

## Installation

You can install the development version of gmeans from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("m-muecke/gmeans")
```

## Usage

The Old Faithful geyser has short and long eruptions, and `gmeans()` finds these two groups without
being told how many to look for:

```{r}
library(gmeans)

km <- gmeans(faithful)
summary(km)
```

## When to use gmeans

Use `gmeans()` when the number of clusters is unknown and the clusters are roughly Gaussian.
The algorithm splits a cluster only when an Anderson-Darling test rejects normality, so the
number of clusters follows from the data and a single significance level rather than from a
grid search over `k`.

The default significance level of `0.0001` follows Hamerly and Elkan (2003) and keeps Gaussian
clusters from being split.
With small clusters of fewer than about 50 points the test has little power to detect a split, so
a larger level such as `0.01` can work better.

## Related work

* [mlr3cluster](https://mlr3cluster.mlr-org.com/): cluster analysis for the mlr3 ecosystem, which provides G-means as the `clust.gmeans` learner.
* [nortest](https://CRAN.R-project.org/package=nortest): R package for testing the composite hypothesis of normality.
